Is the Future of Authenticity All In Our Heads?

### 2017-11-08 authors.tex
Better figures,e tc
changes

\begin{abstract}
Passthoughts, in which a user thinks a secret thought to log in to services or devices, provides two factors of authentication (knowledge and inherence) in a single step. Since its proposal in 2005, passthoughts enjoyed a number of successful empirical studies. In this paper, we renew the promise of passthoughts authentication, reviewing the main challenges that passthoughts must overcome in order to move from the lab to the real world. We propose two studies, which seek different angles at the fundamental questions we pose. We conclude with future opportunities, and challenges, for passthoughts’ broader deployment as a tool for authentication. In doing so, we raise novel possibilities for authentication broadly, such as ``organic passwords'' that change naturally over time, or systems that reject users who are not acting quite ``like themselves.''
\end{abstract}

>>>>

\title{Is the Future of Authenticity In Our Heads?}
\subtitle{Moving Passthoughts From the Lab to the World}

\begin{abstract}
Passthoughts, in which a user thinks a secret thought to log in to services or devices, provides two factors of authentication (knowledge and inherence) in a single step. Since its proposal in 2005, passthoughts enjoyed a number of successful empirical studies. In this paper, we renew the promise of passthoughts authentication, outlining the main challenges that passthoughts must overcome in order to move from the lab to the real world. We propose two studies, which seek different angles at the fundamental questions we pose. Further, we propose it as a fruitful case study for thinking about what authentication can, and should, be expected to do, as it pushes up against questions of what sorts of ``selves'' authentication systems must be tasked with recognizing. Through this discussion, we raise novel possibilities for authentication broadly, such as ``organic passwords'' that change naturally over time, or systems that reject users who are not acting quite ``like themselves.''
\end{abstract}

### 2017-11-08 paper.tex
changes

% Created 2017-09-23 Sat 14:20
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\usepackage{balance}
\date{\today}
\title{Is the Future of Authenticity All In Our Heads?}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity All In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.0.50 (Org mode 9.0.9)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:orgea4abb3}

@@

In addition, computing broadly has undergone many sea changes since passthought's initial proposal in 2005.
Biosensing (the sensing of humans) has proliferated throughout our everyday life, due both to the sensors ubiquitous in smartphones, and to general improvements to infrastructures of data collection and machine learning at scale. 
Is passthoughts more or less appealing in light of other authentication strategies these "sea changes" have made possible?

This paper aims both to re-validate passthought's promise as an authentication paradigm (Section 2), and to 
motivate and contextualize two studies that could begin to address the challenges posed above (Section 5).
While passthoughts is not without risks (Section 6),
it could lead to novel \emph{types} of authentication,
which challenge (or at least destabilize) dominant assumptions about what authentication 
can and should be able to do. (Section 7).

\section{Background}
\label{sec:org5444f31}

Authentication seeks to prove that a user is who they claim to be.
In computer security, authenticators are classified into three types: knowledge factors (e.g., passwords
and PINs), possession factors (e.g., physical tokens, ATM cards), and inherence
factors (e.g., fingerprints and other biometrics).

An ongoing problem in authentication lies in balancing strong security
(i.e., multiple factors of authentication)
with usability.
As an example, major industry players such as Google and
Facebook have strongly encouraged their users to adopt two-factor
authentication (2FA), in which a user enters his or her password (a knowledge factor),
and subsequently receives a code on their cellphone (a posession factor).

@@

\subsection{One-step, multi-factor authentication}
\label{sec:org01cd958}

Since passthought's initial proposal in 2005,
more ubiquitous sensing and computing has enabled a number of other strategies for achieving two factors of authentication in a single step. 
Some work has tested behavioral authentication methods such as keystroke dynamics, or voice. In both cases, the knowledge factor (password or passphrase) and
inherence factor (typing rhythm or speaker's voice) are employed \cite{Monrose1997}.
In contrast, the Nymi band supports one-step two-factor authentication via the inherence
factor (cardiac rhythm that is supposed to be unique to each individual) and the
possession factor (the wearing of the band on the wrist) \cite{Nymi}.
More recent attempts have also used gait (from cellphone accelorometers) to perform authentication \cite{UnifyID2017}.

@@

\subsection{Passthought authentication}
\label{sec:orgf9584e8}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:org8a40167}

@@

The first attempt to merge in-ear EEG with passthought authentication
used a modified consumer grade EEG device with a single electrode, achieving approximately 80 percent authentication accuracy \cite{curranpassthoughts}.
Ongoing work from the same authors investigates the use of custom-fit earbuds with multiple embedded electrodes \ref{fig:earbud}.
Lending credibility to that study's claim that in-ear EEG could one day become feasible in consumer devices,
United Sciences recently announced a consumer "hearable'' (in-ear wearable) called The Aware, which will measure EEG from the ear, among other biometrics.

\section{User attitudes and perceptions}
\label{sec:org4c0e40b}

@@

\subsection{Contending with mind-reading machines}
\label{sec:org850972f}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:org0e23305}

@@

This quote highlights how individuals' life experiences
might shape the way they engage (or refuse to engage) with brain-sensing devices.
In general, this quote and others motivate the need for a rich, qualitative understanding of people's first-hand experiences with brainscanning devices,
as well as data collection,
in order to understand what role BCI applications such as passthoughts could play in day-to-day life.
\section{Diversity and security of passthoughts}
\label{sec:orgf6a3793}

@@

\section{Two studies on passthoughts}
\label{sec:org01b233e}

@@

\subsection{A real-time passthought authenticator}
\label{sec:org550af8d}

@@

\subsubsection{Study protocol}
\label{sec:org91fd7ba}

@@

\subsubsection{The effect of feedback}
\label{sec:org83db71d}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:orgf0b8337}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:org7d812c5}

@@

\subsubsection{Study protocol}
\label{sec:orgd027e0f}

@@

\subsubsection{A more diverse corpus}
\label{sec:org71176c3}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:org0f0eb3a}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how likely or unlikely we are to observe a given passthought in context of a particular person's neural signals
Such analysis between subjects could help shed light how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:org122982a}

@@

\subsubsection{Limitations}
\label{sec:org4b977e1}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org6c5bf79}

@@

\subsection{Privacy}
\label{sec:org09c5036}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org0fca4b0}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:orgc3e42c2}

@@

\section{Further Future Directions}
\label{sec:orgee7020f}

@@

\subsection{Continuous authentication}
\label{sec:orgad27611}

@@

\subsection{Organic passwords}
\label{sec:orgec1b8eb}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Authentication and the self}
\label{sec:org8dbd4d5}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as we are aware, novel, 
and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this developing method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:orgf6d67e3}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:orgd23d062}

@@

Again, these opportunities must strike a balance with the risks of individual users' privacy and security.
Violating user privacy by revealing EEG data, even to researchers, could undermine any chance of wider BCI adoption in the long-term.
Striking this balance will require a deeper understanding of the statistical properties of signals. 
How much data will users really need to give up? 
What counts as an ``anomalous'' reading?
Answers to these questions could themselves inform neuroscientific inquiry.
This balance will also require a deeper understanding of individuals' attitudes about the meaning of such signals,
and how private people believe them to be.

\section{Conclusion}
\label{sec:org056f202}

>>>>

% Created 2017-11-08 Wed 10:17
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\usepackage{balance}
\date{\today}
\title{Is the Future of Authenticity All In Our Heads?}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity All In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.0.50 (Org mode 9.1.2)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org643f242}

@@

In addition, computing broadly has undergone many sea changes since
passthought's initial proposal in 2005. Biosensing (the sensing of humans) has
proliferated throughout our everyday life, due both to the sensors ubiquitous in
smartphones, and to general improvements to infrastructures of data collection
and machine learning at scale. Is passthoughts more or less appealing in light
of other authentication strategies these ``sea changes'' have made possible?

This paper aims both to re-validate passthought's promise as an authentication
paradigm (Section 2), and to motivate and contextualize two studies that could
begin to address the challenges posed above (Section 5). While passthoughts is
not without risks (Section 6), it could lead to novel \emph{types} of authentication,
which challenge (or at least destabilize) dominant assumptions about what
authentication can and should be able to do. (Section 7). We argue
that passthoughts is a useful case for thinking about authentication, as it
raises interesting philosophical questions about what the self \emph{is} such that
authentication systems might seek to detect it.

\section{Background}
\label{sec:orgc2af2a9}

Authentication seeks to prove that a user is who they claim to be. In computer
security, authenticators are classified into three types: knowledge factors
(e.g., passwords and PINs), possession factors (e.g., physical tokens, ATM
cards), and inherence factors (e.g., fingerprints and other biometrics).
Passwords, which rely only on a knowledge factor, offer many benefits
\cite{Bonneau2012}, including their capability for allowing \emph{authorization}
(granting use of a system). However, for authentication (verifying an
individual), passwords are easy to compromise, and easy to forget. These
shortcomings motivate the overlay of multiple factors of authentication.

An ongoing problem in authentication lies in balancing strong security (i.e.,
multiple factors) with usability. As an example, major industry players such as
Google and Facebook have strongly encouraged their users to adopt two-factor
authentication (2FA), in which a user enters his or her password (a knowledge
factor), and subsequently receives a code on their cellphone (a posession
factor).

@@

\subsection{One-step, multi-factor authentication}
\label{sec:org0b6dedd}

Since passthoughts' initial proposal in 2005, more ubiquitous sensing and
computing has enabled a number of other strategies for achieving two factors of
authentication in a single step. Some work has tested behavioral authentication
methods such as keystroke dynamics, or voice. In both cases, the knowledge
factor (password or passphrase) and inherence factor (typing rhythm or speaker's
voice) are employed \cite{Monrose1997}. In contrast, the Nymi band supports
one-step two-factor authentication via the inherence factor (cardiac rhythm that
is supposed to be unique to each individual) and the possession factor (the
wearing of the band on the wrist) \cite{Nymi}. More recent attempts have also used
gait (from cellphone accelorometers) to perform authentication \cite{UnifyID2017}.

@@

\subsection{Passthought authentication}
\label{sec:orgc42e0c7}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:orgc61af97}

@@

The first attempt to merge in-ear EEG with passthought authentication
used a modified consumer grade EEG device with a single electrode, achieving approximately 80 percent authentication accuracy \cite{curranpassthoughts}.
Ongoing work from the same authors investigates the use of custom-fit earbuds with multiple embedded electrodes \ref{fig:earbud}.
Lending credibility to that study's claim that in-ear EEG could one day become feasible in consumer devices,
United Sciences recently announced a consumer ``hearable'' (in-ear wearable) called The Aware, which will measure EEG from the ear, among other biometrics.

\section{User attitudes and perceptions}
\label{sec:org244a326}

@@

\subsection{Contending with mind-reading machines}
\label{sec:orgf9391d3}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:org230b3b8}

@@

This quote highlights how individuals' life experiences
might shape the way they engage (or refuse to engage) with brain-sensing devices.
In general, this quote and others motivate the need for a rich, qualitative understanding of people's first-hand experiences with brainscanning devices,
as well as data collection,
in order to understand what role BCI applications such as passthoughts could play in day-to-day life.
\section{Diversity and security of passthoughts}
\label{sec:org0027258}

@@

\section{Two studies on passthoughts}
\label{sec:org883ae3d}

@@

\subsection{A real-time passthought authenticator}
\label{sec:org509efae}

@@

\subsubsection{Study protocol}
\label{sec:orgf3dd9d1}

@@

\subsubsection{The effect of feedback}
\label{sec:orged55079}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:orgb9eab74}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:org2a09ca4}

@@

\subsubsection{Study protocol}
\label{sec:org38209c6}

@@

\subsubsection{A more diverse corpus}
\label{sec:org848ad1f}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:orgb3c246c}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how likely or unlikely we are to observe a given passthought in context of a particular person's neural signals
Such analysis between subjects could help shed light how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:org88bd963}

@@

This deployment would also provide a number of opportunities for better
understanding how passthoughts compares to other modes of authentication.
Subjects would need to set up the device themselves, and could compare the
device's use with their existing password practices. Subjects might also need to
share passthoughts, which could cause unforeseen complications, especially for
persons with disabilities \cite{Dosono2015}.

At-home deployments could also allow for investigation of subjects' capacities
for remembering their passthoughts. A study could ask subjects to remember a
number of passthoughts, to see how many passthoughts can be remembered (and how
many different passthoughts an authentication system could detect). The study
may also ask subjects to change their passthoughts, to see how easy or difficult
it is for users to adapt if their pasthought needs to be reset.

\subsubsection{Limitations}
\label{sec:org42611a6}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:orgb8a670d}

@@

\subsection{Privacy}
\label{sec:org3367374}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org1e5fa6b}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:org639a96b}

@@

\section{Discussion}
\label{sec:orge41f8ae}

@@

\subsection{Continuous authentication}
\label{sec:orgbac1c33}

@@

\subsection{Organic passwords}
\label{sec:org252a6e3}

If EEG signals are nonstationary (changing over time), passthoughts will require
continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}. This
feature of BCIs could have an unexpected benefit to security. If an individual's
expression of their passthought in EEG is always changing, passthoughts
themselves are effectively evergreen, automatically replaced or updated by
nature of the authentication paradigm. This feature could improve security, as
an attacker able to compromise a passthought's EEG signature may not be able to
log into the system in a few weeks time, unless they are able to realistically
mutate the signal over authentication attempts. This feature of EEG also gives
passthoughts a possible advantage over other methods for behavioral
authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may
change more slowly for individuals, if they change at all. Future work should
investigate this claim, perhaps using a longitudinal corpus such as the one
described above.

\subsection{Attacks on passthoughts}
\label{sec:orgbd5fb9d}

Passthoughts has yet to be subjected to rigorous threat analysis. One obvious
threat is a simple replay attack, in which noise is added to recorded
(intercepted) authentication attempts in order to trick the classifier into
acceptance. Passthoughts' susceptibility to such an attack effectively reduces
to the security of its components (the EEG scanner, the transmission device,
etc). Thus, future work should examine the possible new attack vectors
introduced by passthoughts' equipment, as these attack vectors
could pave the way for attacks such as replays.

A more interesting mode of attack could involved a principled search through the
space of EEG signals. For example, a medical corpus of EEG signals might provide
enough information about the distribution of EEG readings such that an attacker
can generate convincing EEG signals. How these EEG signals might be used to
discover the underlying properties of a given subject's space of passthoughts
remains unknown. In either case, such a process could be viewed as a sort of
dictionary attack, where existing knowledge about the space of passthoughts
generally is used to compromise (or reduce the search space for) a particular
passthought. Future work should employ various methods for synthesizing EEG
signals in order to gain a greater appreciation for how susceptible passthoughts
might be to such dictionary-style attacks.

\subsection{Authentication and the self}
\label{sec:org297fe9f}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions.
How does such a system conform to accepted definitions of a ``person''? Who is a
person to make decisions for their future self? What are possible vectors for
abuse? In any case, this property of an authentication is, as far as we are
aware, novel, and should be considered as we learn more about the strengths,
weaknesses, and particular affordances of this developing method for
authentication. Future work might compare such scenarios with existing
definitions of personhood and consent in legal systems, which may provide
guidance (or, at minimum, a reference point) for such discussions.
\subsection{Passthoughts by any other sensor?}
\label{sec:org8f527c9}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:org0ff39f4}

@@

Again, these opportunities must strike a balance with the risks of individual users' privacy and security.
Violating user privacy by revealing EEG data, even to researchers, could undermine any chance of wider BCI adoption in the long-term.
Striking this balance will require a deeper understanding of the statistical properties of signals. 
How much data will users really need to give up? 
What counts as an ``anomalous'' reading?
Answers to these questions could themselves inform neuroscientific inquiry.
This balance will also require a deeper understanding of individuals' attitudes about the meaning of such signals,
and how private people believe them to be.
\section{Conclusion}
\label{sec:org50f75b6}
